04-analisis

### 2024-05-27 03-methods.qmd
Actualizacion

# Data, variables and method

## Data

## 🗒️ Answer

@@

## Variables

@@

**Justification of Inequality**: The dependent variables in this study are market justice preferences. This construct is measured through three variables addressing the degree of justification regarding whether access to social services in pensions, education and health should be income conditional. Justification of inequality in health is measured by the item "Is it fair in Chile that people with higher incomes can access better healthcare than those with lower incomes,", and the same is asked for education and pensions. In all cases, respondents declare their preferences on a Likert scale ranging from "completely disagree" (1) to "completely agree" (4). Additionally, we include a summarized indicator of preferences for market justice (REF), measured by an average index across these items (α = 0.86), with values ranging from 1 to 4, where higher values represent greater preferences for market justice (see @tbl-desc-dependientes). We analyzed these items independently as well as by constructing and average index called "market justice preferences".

@@

The analysis includes several individual control variables that have been identified as influential in recent research. These variables encompass family socioeconomic status - by the highest level of parental education -, the number of books in the household, and an index measuring access to technologies. @tbl-desc-independent shows the independent variables used, their response categories and their frequencies.

@@

df<-dfSummary(independientes,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(30,10,10,10))

@@

This study examines four variables at the school level. Firstly, we assess the administrative dependence of schools, categorized as "Public," "Subsidized Private," or "Private." Secondly, we consider the socioeconomic classification of schools, determined by the Ministry of Education and measured on an ordinal scale with five categories ranging from "low" (1) to "high" (5). Thirdly, we evaluate the performance level of schools in SIMCE standardized tests, categorized as "low," "medium," or "high". SIMCE (System of Measurement of Educational Quality) is a national test in areas such as language and mathematics which is applied yearly in different educational levels (REF). Lastly, we analyze the proportion of parents with university or postgraduate education level at each school. The variables, response categories, and their frequencies are detailed in @tbl-desc-school. {@kevin arreglar N L2 tabla}

@@

df<-dfSummary(school,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(30,10,10,10))

@@

## Methods

>>>>

## Data, variables and method

### Data

@@

### Variables

@@

**Justification of Inequality**: The dependent variables in this study are market justice preferences. This construct is measured through three variables that address the degree of justification regarding whether access to social services in pensions, education, and health should be income conditional. Justification of inequality in health is measured by the item: "Is it fair in Chile that people with higher incomes can access better healthcare than those with lower incomes?" The same question is asked for education and pensions. In all cases, respondents indicate their preferences on a Likert scale ranging from "completely disagree" (1) to "completely agree" (4). Additionally, we include a summarized indicator of "market justice preferences", measured by an average index across these items (α = 0.86), with values ranging from 1 to 4, where higher values represent greater preferences for market justice (see @tbl-desc-dependientes). We analyzed these items independently as well as by the average index.

@@

**Family Socioeconomic Status**: The socioeconomic status of students' families is measured using two indicators. First, the highest educational level attained by the parents, with categories: "8th grade or less," "Secondary education," "Technical higher education," "University or postgraduate," and "No response." The inclusion of the "No response" category is due to its high frequency in the data; omitting it could obscure relevant associations. Second, the number of books in the household is used, categorized as "Less than 25" and "More than 25."

@tbl-desc-independent shows the independent variables used, their response categories and their frequencies.

@@

df<-dfSummary(independientes[,-8],
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(30,10,10,10))

@@

This study focuses on two school-level characteristics: socioeconomic status and academic performance. Socioeconomic status is assessed using the Ministry of Education's classification, measured as an ordinal item with five categories ranging from "low" (1) to "high" (5). Academic performance is measured using the school's results in the SIMCE (System of Measurement of Educational Quality) standardized tests, which are administered yearly at different educational levels in areas such as language and mathematics. These results are categorized as "low," "medium," and "high." The items used, response categories, and their frequencies are detailed in @tbl-desc-school.

@@

df<-dfSummary(school[,c(2,4)],
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(30,10,10,10))

@@

***Controls***

A series of individual control variables are included. At the individual level, an index of access to technology is constructed based on the number of computers, tablets, and cell phones in the household, as well as internet connectivity. At the contextual level, we use the administrative dependence of schools, classified as "Public," "Subsidized Private," or "Private," and the proportion of parents with university or postgraduate education at each school.

### Methods

### 2024-05-27 04-analisis.qmd
revisada introduccion y antecedentes, temas menores formato
Actualizacion

---
author: "Equipo Edumer"
format: pdf
editor: visual
---

@@

```{r packages, echo=FALSE}
pacman::p_load(tidyverse,
               ggpubr, 
               lme4, 
               ordinal,
               sjPlot, 
               texreg,
               kableExtra)

options(scipen=999)
rm(list = ls())

@@

@@

@@

The results of the cumulative link mixed models for the ordinal dependent variables of justice in differential access to education, health, and pensions by income are presented in @fig-odds. This figure illustrates the effect of meritocratic variables on society and school as independent variables; however, the models incorporate all variables utilized in the study for each of the three dependent variables separately (complete models are available in the [Appendix](#appendix)).

```{r echo=FALSE}

data_pension <- data_rec %>% dplyr::select(-just_educ, -just_salud, -redistribucion)
data_pension <- na.omit(data_pension)
data_pension$just_pension_fact <- as_factor(data_pension$just_pension)
reg5_pension <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_pension)
data_salud <- data_rec %>% dplyr::select(-just_pension, -just_educ, -redistribucion)
data_salud <- na.omit(data_salud)
data_salud$just_salud_fact <- as_factor(data_salud$just_salud)
reg5_salud <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_salud)
data_educ <- data_rec %>% dplyr::select(-just_pension, -just_salud, -redistribucion)
data_educ <- na.omit(data_educ)
data_educ$just_educ_fact <- as_factor(data_educ$just_educ)
reg5_educ <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + cod_grupo + simce + (1 | mrbd), data=data_educ)
```

```{r echo=FALSE}
#| fig-cap: "Odds-ratios of justification of pensions, education and health by social and school meritocracy"
#| label: fig-odds
#| fig-asp: 0.8

plot_models(reg5_salud, reg5_educ, reg5_pension,  
            rm.terms = c("1|2", "2|3", "3|4", "educacion_recEd Media", "educacion_recEd Tecnica", "educacion_recUniversidad o posgrado", "educacion_recNs/Nr", "libros_recMás de 25", "acc_tec", "internet_rec2", "internet_recNs/Nr", "mean_educ", "cod_depe2Part. subvencionado", "cod_depe2Part. privado", "cod_grupoMedio bajo", "cod_grupoMedio", "cod_grupoMedio alto", "cod_grupoAlto", "simceMedio", "simceAlto"), #quitar terminos
            title = "", #quitar titulo
            show.values = TRUE, #mostrar valor de efectos
            dot.size = 1.3, #tamaño circulos
            line.size = 1, #tamaño CI
            value.size = 3.5, #tamaño valor efectoss
            spacing = .7, #espacio entre efectos
            grid = TRUE, # separa en un plot por modelo
            vline.color = "firebrick", # linea roja en punto neutro (1)
            axis.labels = rev(c("School talent", 
                            "School effort", 
                            "Social talent",
                            "Social effort", 
                            "Deservingness"
                            )), #con rev porque automatico los tira en otro orden
            show.legend = FALSE,
            m.labels = c("Health", "Education","Pension"),
            colors = gray(level = 0.2))  # variables dependientes

@@

@fig-odds demonstrates a similar trend for the three dependent variables of justice in differential access to pensions, education, and health, with all effects being statistically significant. In the context of school meritocracy, the effects are mixed: as perceived school talent increases, the justification for differentiated access to these services increases; conversely, as perceived school effort increases, the justification decreases, holding all other variables constant. Regarding the variables of social meritocracy, the three variables of talent, effort, and deservingness show that as these increase, the justification for differentiated access to pensions, education, and health also increases. Overall, these results provide partial support for our first hypothesis (H1a), which posited that students who perceive higher levels of meritocracy in school would be more likely to justify differential access to pensions, health, and education. This is evidenced by the positive effect of perceiving talent as important for achieving good grades. However, the results strongly support our second hypothesis (H1b), as a higher perception of societal meritocracy among students tends to increase the justification for market-based justice in these social services.

>>>>

---
author: "Equipo Edumer"
editor: visual
---

@@

```{r packages, echo=FALSE}
pacman::p_load(tidyverse,
               ggpubr, 
               lme4, 
               ordinal,
               sjPlot, 
               texreg,
               kableExtra,
               easystats)

options(scipen=999)
rm(list = ls())
```

@@

odds_fun <- function(modelos) {
   or <- texreg::extract(modelos)
   or@coef <- exp(or@coef)
   return(or)
}

@@

Likert plots

@@

The results of the cumulative link mixed models for the ordinal dependent variables of justice in differential access to education, health, and pensions by income are presented in @tbl-ordinal-reg. This table shows the effects of individual and contextual-level independent variables related to our hypothesis; however, the models also incorporate the control variables utilized in the study for each of the three dependent variables separately (complete models are available in the [Appendix](#appendix)).

data_pension <- data_rec %>% 
    dplyr::select(-just_educ, -just_salud, -redistribucion)
data_pension <- na.omit(data_pension)
data_pension$just_pension_fact <- as_factor(data_pension$just_pension)

m0_p <- clmm(just_pension_fact ~ 1 + (1 | mrbd), data=data_pension)

performance::icc(m0_p, by_group = T)

m1_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_pension)

m2_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_pension)

m3_p <- clmm(just_pension_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
    cod_grupo + simce + (1 | mrbd), data=data_pension)

tbls <- list(m1_p, m2_p, m3_p)

modelos_odd_pension <- map(tbls, odds_fun)

data_salud <- data_rec %>% 
    dplyr::select(-just_pension, -just_educ, -redistribucion)
data_salud <- na.omit(data_salud)
data_salud$just_salud_fact <- as_factor(data_salud$just_salud)

m0_h <- clmm(just_salud_fact ~ 1 + (1 | mrbd), data=data_salud)

performance::icc(m0_h, by_group = T)

m1_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_salud)

m2_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_salud)

m3_h <- clmm(just_salud_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
    cod_grupo + simce + (1 | mrbd), data=data_salud)

tbls <- list(m1_h, m2_h, m3_h)

modelos_odd_health <- map(tbls, odds_fun)

data_educ <- data_rec %>% 
    dplyr::select(-just_pension, -just_salud, -redistribucion)
data_educ <- na.omit(data_educ)
data_educ$just_educ_fact <- as_factor(data_educ$just_educ)

m0_e <- clmm(just_educ_fact ~ 1 + (1 | mrbd), data=data_educ)

performance::icc(m0_e, by_group = T)

m1_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + (1 | mrbd), data=data_educ)

m2_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
    libros_rec + acc_tec + internet_rec + (1 | mrbd), data=data_educ)

m3_e <- clmm(just_educ_fact ~ 1 + inteligencia_esc + esfuerzo_esc + 
    inteligencia_soc + esfuerzo_soc + merito_soc + educacion_rec + 
    libros_rec + acc_tec + internet_rec + mean_educ + cod_depe2 + 
    cod_grupo + simce + (1 | mrbd), data=data_educ)

tbls <- list(m1_e, m2_e, m3_e)

modelos_odd_educ <- map(tbls, odds_fun)

@@

```{r echo=FALSE, results='asis'}
#| tbl-cap: "Cumulative link multilevel models of differential acces justificacion of education, health and pensions"
#| label: tbl-ordinal-reg

ccoef <- list(
  inteligencia_esc = "School talent",
  esfuerzo_esc = "School effort",
  inteligencia_soc = "Social talent",
  esfuerzo_soc = "Social effort",
  merito_soc = "Deservingness",
  "educacion_recEd Media" = "Secondary",
  "educacion_recEd Tecnica" = "Higher tec.",
  "educacion_recUniversidad o posgrado" = "University or posgraduate",
  "educacion_recNs/Nr" = "Missing",
  mean_educ = "Prop. university level at school",
  "cod_grupoMedio bajo" = "SES Medium low",
  cod_grupoMedio = "SES Medium",
  "cod_grupoMedio alto" = "SES Medium high",
  cod_grupoAlto = "SES High",
  simceMedio = "Simce Medium",
  simceAlto = "Simce High"
)

texreg::htmlreg(l = list(modelos_odd_educ[[2]],
                         modelos_odd_educ[[3]],
                         modelos_odd_health[[2]],
                         modelos_odd_health[[3]],
                         modelos_odd_pension[[2]],
                         modelos_odd_pension[[3]]),   
                  stars = c(0.05, 0.01, 0.001),
                  custom.note = "Note: Coefficients models in odds-ratio. Control variables are included. Standard errors in parentheses. %stars",
                  custom.header = list("Education" = 1:2, "Health" = 3:4, "Pensions" = 5:6),
                  custom.model.names = rep(c("Model 1", "Model 2"), 3),
                  custom.coef.map = ccoef,
                  leading.zero = T,
                  scalebox = 0.60,
                  groups = list("Parental education (Ref.= 8th grade or less)" = 6:9,
                                "Socioeconomic level (Ref.= Low)" = 11:14,
                                "Achievement score (Ref.= Low)" = 15:16),
                 custom.gof.rows = list("Controls"=rep("Yes",6)))

@tbl-ordinal-reg demonstrates a similar trend for the three dependent variables of justice in differential access to pensions, education, and health, with all effects being statistically significant. In the context of school meritocracy, the effects are mixed: as perceived school talent increases, the justification for differentiated access to these services increases; conversely, as perceived school effort increases, the justification decreases, holding all other variables constant. Regarding the variables of social meritocracy, the three variables of talent, effort, and deservingness show that as these increase, the justification for differentiated access to pensions, education, and health also increases. Overall, these results provide partial support for our first hypothesis (H1a), which posited that students who perceive higher levels of meritocracy in school would be more likely to justify differential access to pensions, health, and education. This is evidenced by the positive effect of perceiving talent as important for achieving good grades. However, the results strongly support our second hypothesis (H1b), as a higher perception of societal meritocracy among students tends to increase the justification for market-based justice in these social services. CAMBIAR
